---
title: "baseR_markdown"
author: "Glory"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Reading data practice
```{r}
file <- readRDS("my_day2matrix.rds")
file
```

### Question 3 Adding rows and columns

```{r}
rowSums <- rowSums(file, na.rm = TRUE)
rowMeans <- rowMeans(file, na.rm = TRUE)
file_1 <- cbind(file, rowSums, rowMeans)
file_1
```

### Question 4 loading data-ToothGrowth

```{r}
data("ToothGrowth")
```

### Question 5 More info on dataset

```{r}
summary(ToothGrowth)

```

### Columns in dataset

```{r}
head(ToothGrowth)
```

### Mean toothlength

```{r}
mean(ToothGrowth$len)
```

### Minimum and maximum legnth

```{r}
max(ToothGrowth$len)
min(ToothGrowth$len)
```

### calculate the row sum and Mean of dataset

```{r}
rowSums(ToothGrowth)
rowMeans(ToothGrowth)
```
cannot because of non numeric data

### Creating a function

```{r}
triangle <- function(a, b) {
               hypotenuse <- sqrt(a^2 + b^2)
                     return(hypotenuse)
}

triangle(2,9)
```

### Using apply function

```{r}
vec <- c(10:20)
square <- function(y) {
  sq <- y^2
  return(sq)
}
square(4)

# use lapply
lapply(vec, square)
# use sapply
sapply(vec, square)
```

### using apply function on list

```{r}
my_list <- list(c(1:10), c(2:4), vec, c(T,F,T,T,F))
lapply(my_list, sum)
sapply(my_list, sum)
```
### using rep with apply function

```{r}
new_list <- sapply(my_list, rep, each = 3)
new_list
```

