---
title: "Example code for ggplot2"
author: "Kevin Rue-Albrecht"
date: "03/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(patchwork)
library(gridExtra)
```

# Exercise

Use `library(ggplot2)` to load the package.

```{r}
library(ggplot2)
```

## Specify the dataset.

What does `ggplot(diamonds)` do?

```{r}
ggplot(diamonds)
```

> Answer: It plotted a blank graph, because nothing was specified in the command.

## Add the aesthetics.

What does `ggplot(diamonds, aes(x = carat, y = price))` do?

```{r}
ggplot(diamonds, aes(x = carat, y = price))
```

> Answer:It plotted the dimensions for the y and x axis, but the type of graph was yet to be specified

## Add geometric objects

- Add data points showing `carat` on the x-axis and `price` on the y-axis.

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point()
  
```

- Color data points by `cut`.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point()
  
```

- Add a smoothed mean trend line.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point() + 
  geom_smooth()

```

- Assign that last plot to an object called `obds_diamonds`.

```{r}
obds_diamonds <- ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point() + 
  geom_smooth()

```


# Exercise

Predict the difference between these two plots

### Plot 1

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price, colour = cut)) +
  geom_point() +
  geom_smooth()
```

### Plot 2

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(colour = cut)) +
  geom_smooth()
```

```{r}
install.packages("ggpubr")
```

> Answer:In plot 1 the colour was specified globally which resulted in the generation of mean lines for all diamond cuts
However, the colour by cut was specified at the geom_point in plot 2 theerfore only one mean line was generated

# Exercise

## Trend lines

Using the `ChickWeight` data set:

- Create a scatter plot of weight (y-axis) over time (x-axis).

```{r}
data("ChickWeight")
ggplot(ChickWeight, aes(x = Time, y = weight) ) +
  geom_point()

```

- Color by diet.
  (Copy-paste and extend the code chunk above.)

```{r}
data("ChickWeight")
ggplot(ChickWeight, aes(x = Time, y = weight, color = Diet) ) +
  geom_point()
```

- Add a linear mean trend line for each diet.
  (Copy-paste and extend the code chunk above.)

```{r}
data("ChickWeight")
ggplot(ChickWeight, aes(x = Time, y = weight, color = Diet) ) +
  geom_point() +
  geom_smooth(method = "lm")
```

At this point you should be able to visually identify
the diet that leads to the highest mean increase in weight.

> Answer: purple

- Facet a ribbon of sub-plots, one per diet.
  (Copy-paste and extend the code chunk above.)

```{r}
data("ChickWeight")
ggplot(ChickWeight, aes(x = Time, y = weight, color = Diet) ) +
  geom_point() +
  facet_wrap(~Diet, nrow = 2)
```

- Assign that last plot to an object called `obds_chickweight`.

```{r}
data("ChickWeight")
obds_chickweight <- ggplot(ChickWeight, aes(x = Time, y = weight, color = Diet) ) +
  geom_point() +
  facet_wrap(~Diet, nrow = 2)
```

# Exercise

## Bar plot

- Load the `ggplot2::msleep` data set.

```{r}
ggplot2::msleep
```

- Draw a bar plot of number of observations (i.e., rows) for each taxonomic order
  (i.e, one plot and one bar per taxonomic order).

```{r}
ggplot(msleep, aes(x = order)) +
geom_bar()
```

- Change the angle and font size of the text for the x-axis ticks (not the axis titles).
  Justify the text of those x-axis ticks as right-aligned.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(msleep, aes(x=order)) +
geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1, vjust = 1))

```

- Change the value and font size of the title for both x and y axes.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(msleep, aes(x=order)) +
geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1, vjust = 1))+
  labs(title = "msleep plot", x = "order", y = "frequency") +
  theme(axis.title = element_text(size = 14, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```

- Fill each bar with colors, proportionally to the count of each genus.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(msleep, aes(x=order)) +
geom_bar(aes(fill = genus)) + 
  theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1, vjust = 1))+
  labs(title = "msleep plot", x = "order", y = "frequency") +
  theme(axis.title = element_text(size = 14, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```

From this point onwards, you may need to iteratively resize the text of the ticks and axes for readability.

- Reduce the legend key size.
  (Recommendation: use `unit(2, "mm")`).
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(msleep, aes(x=order)) +
geom_bar(aes(fill = genus)) + 
  theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1, vjust = 1))+
  labs(title = "msleep plot", x = "order", y = "frequency") +
  theme(axis.title = element_text(size = 14, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  theme(legend.key.size = unit(1, "mm"))
```

- Force the legend to be display in 3 columns.
  (Recommendation: use `guide_legend(...)`).
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(msleep, aes(x=order)) +
geom_bar(aes(fill = genus)) + 
  theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1, vjust = 1))+
  labs(title = "msleep plot", x = "order", y = "frequency") +
  theme(axis.title = element_text(size = 14, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  guides(fill = guide_legend(ncol=3)) +
  theme(legend.key.size = unit(1, "mm"))
```

- Add a contour of thin black lines to the bars.

```{r}
ggplot(msleep, aes(x=order)) +
geom_bar(aes(fill = genus), color="black", linewidth=0.5) + 
  theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1, vjust = 1))+
  labs(title = "msleep plot", x = "order", y = "frequency") +
  theme(axis.title = element_text(size = 14, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  guides(fill = guide_legend(ncol=3)) +
  theme(legend.key.size = unit(1, "mm"))
```

- Assign that last plot to an object called `obds_msleep`.

```{r}
obds_msleep <- ggplot(msleep, aes(x=order)) +
geom_bar(aes(fill = genus), color="black", linewidth=0.5) + 
  theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1, vjust = 1))+
  labs(title = "msleep plot", x = "order", y = "frequency") +
  theme(axis.title = element_text(size = 14, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  guides(fill = guide_legend(ncol=3)) +
  theme(legend.key.size = unit(1, "mm"))
```

# Exercise

## Plotting grid

- Collate the plots that we assigned to objects through the day, as a single plot.

  + Plots: `obds_diamonds`, `obds_chickweight`, `obds_msleep`.
  
  + Methods: `cowplot::plot_grid()`, `patchwork`, `gridExtra::grid.arrange()`.

Using `cowplot`.

```{r, fig.height=6, fig.width=12}
library(cowplot)
first_row <- cowplot::plot_grid(obds_diamonds, obds_chickweight, nrow = 1
                                
)
second_row <- cowplot::plot_grid(obds_msleep, nrow = 1
                                 
)
super_plot <- cowplot::plot_grid(first_row, second_row, labels = c("A", "B", "C",  nrow = 2)
                                 
)
super_plot
```

Using `patchwork`.

```{r, fig.height=9, fig.width=12}
library(patchwork)
plot_patchwork <- (obds_diamonds + obds_chickweight) / (obds_msleep)
```

Using `gridExtra`.

```{r, fig.height=9, fig.width=12}
library(gridExtra)
plot_gridExtra <- gridExtra::grid.arrange(obds_diamonds, obds_chickweight, obds_msleep, ncol = 2, nrow = 2)
```

- Export the new plot in a PDF file, and open it in a PDF viewer (e.g. Adobe Acrobat Reader DC).

You will likely need a few attempts to fine-tune the width and height of the output file.

```{r}
ggsave( 
  "combined_graph.jpeg",
  plot = plot_patchwork
  )
```

# Exercise

## Pair programming

- Explore the data set `ggplot2::mpg` and generate the most informative plot that you can!

```{r}




```
